# MME: No sexID strat


# Het, gay/lesbian, bi population, sexual activity stratification only

In this model, we do not stratify by sexual identity or restrict to heterosexual sexual identity. The analysis procedure is exactly the same as for the heterosexual-only population.

## Estimate the number of new partnerships with each sex

```{r}
het_partners <- all_sex %>%
    group_by(r_sex, r_sexact) %>%
    summarise(partners = weighted.mean(hetnonew, w = total_wt)) %>%
    mutate(rp_sex = ifelse(r_sex == "m", 'w', 'm')) %>%
    rowwise() %>%
    mutate(prop = get_proportion(all_sex,
                                 q_sex = r_sex,
                                 q_sexact = r_sexact),
           n_partners = partners * prop) %>%
    ungroup()

sam_partners <- all_sex %>%
    group_by(r_sex, r_sexact) %>%
    summarise(partners = weighted.mean(samnonew, w = total_wt)) %>%
    mutate(rp_sex = ifelse(r_sex == "m", 'm', 'w')) %>%
    rowwise() %>%
    mutate(prop = get_proportion(all_sex,
                                 q_sex = r_sex,
                                 q_sexact = r_sexact),
           n_partners = partners * prop) %>%
    ungroup()

sexid_nostrat_rep <- rbind(het_partners, sam_partners) %>%
  arrange(r_sex, r_sexact)
  
format_table(sexid_nostrat_rep)
```

Note that the sum of the `prop` column is now 2:

```{r}
sum(sexid_nostrat_rep$prop)
```

That's because each demographic group has two entries in the table: one for partnerships with the same sex and one for the opposite sex. This format makes it easier to do the balancing, as above.

Next, let's distribute the partnerships across activity levels.

## Distribute partnerships

```{r}
sexid_nostrat_offered <-  sexid_nostrat_rep %>%
  group_by(rp_sex, r_sex) %>%
  mutate(prop_to_r = n_partners / sum(n_partners)) %>%
  select(rp_sex,
         r_sex,
         rp_sexact = r_sexact,
         prop_to_r) %>%
  ungroup() %>%
  arrange(rp_sex, r_sex)
format_table(sexid_nostrat_offered)
```

Here, the percentages refer to the proportion of partnerships with people of sex `rp_sex` that are offered to `r_sex` by people with sex `rp_sex` and activity `rp_sexact`. For example, high-activity men offer about 91% percent of male partnerships to men, and high-activity women offer about 66% of female partnerships to men.

Now, we join the offered table and the reported table.

```{r}
sexid_nostrat_joined <- left_join(sexid_nostrat_offered, sexid_nostrat_rep, by = c("r_sex", "rp_sex"))
format_table(sexid_nostrat_joined)
```

The multiplication gives us the distributed parterships. Then, we define a single variable that describes the two demographic variables (sex and SA group):

```{r}
partner_dist_sexid_nostrat <- sexid_nostrat_joined %>%
  mutate(d_partners = partners * prop_to_r,
         r_demo = paste(r_sex, r_sexact, sep="_"),
         rp_demo = paste(rp_sex, rp_sexact, sep="_")) %>%
  select(r_demo, rp_demo, d_partners, prop)
format_table(partner_dist_sexid_nostrat)
```

So, men with high sexual activity reported `r format(sexid_nostrat_rep$partners[1], digits = 3)` partners with women. In the distributed table, these partners are distributed across women with high sexual activity and women with low sexual activity.

```{r}
men_high_sa <- partner_dist_sexid_nostrat %>%
  filter(r_demo == "m_high" & (rp_demo == "w_high" | rp_demo == "w_low"))
format_table(men_high_sa)
sum(men_high_sa$d_partners)
```

Similarly, high-SA men reported `r format(sexid_nostrat_rep$partners[2], digits = 3)` with men on average, which are distributed across men by sexual activity:

```{r}
men_high_sa <- partner_dist_sexid_nostrat %>%
  filter(r_demo == "m_high" & (rp_demo == "m_high" | rp_demo == "m_low"))
format_table(men_high_sa)
sum(men_high_sa$d_partners)
```

## Balancing

Join the distributed data frame with itself:

```{r}
sexid_nostrat_bidi <- partner_dist_sexid_nostrat %>%
    # do a self-join t calculate partners from perspective of rp
    inner_join(partner_dist_sexid_nostrat,
               by = c("r_demo" = "rp_demo",
                      "rp_demo" = "r_demo"),
               suffix = c('.r', '.rp'))
sexid_nostrat_bidi %>%
  format_table()
```


```{r}
theta <- 0.5
b_sexid_nostrat_bidi <- sexid_nostrat_bidi %>%
    mutate(np_r = d_partners.r * prop.r,
           np_rp = d_partners.rp * prop.rp) %>%
    mutate(imbalance = np_r / np_rp,
           corrected_r = d_partners.r / imbalance^(1 - theta),
           corrected_rp = d_partners.rp * imbalance^theta,
           cnr = corrected_r * prop.r,
           cnrp = corrected_rp * prop.rp) %>%
    select(r_demo,
           rp_demo,
           prop.r,
           np_r,
           prop.rp,
           np_rp,
           d_partners.r,
           corrected_r,
           d_partners.rp,
           corrected_rp,
           imbalance,
           cnr,
           cnrp) %>%
  arrange(r_demo, rp_demo)

format_table(b_sexid_nostrat_bidi)
all(with(b_sexid_nostrat_bidi, abs(cnr - cnrp) < .Machine$double.eps))
```

Balanced!

### Compare

```{r}
library(stringr)
sexid_nos_by_grp <- b_sexid_nostrat_bidi %>%
  mutate(rp_sex = str_split(rp_demo, "_", simplify = TRUE)[, 1]) %>%
  group_by(r_demo, rp_sex) %>%
  summarize(bal_p_pc = sum(corrected_r))

sexid_nos_compare <- sexid_nostrat_rep %>%
  mutate(r_demo = paste(r_sex, r_sexact, sep = "_")) %>%
  select(r_demo, rp_sex, orig_p_pc = partners) %>%
  left_join(sexid_nos_by_grp, by = c("r_demo", "rp_sex"))
format_table(sexid_nos_compare)
```

```{r}
# plot with nas
library(ggplot2)
plot_mixing_matrix(b_sexid_nostrat_bidi)
```

# Sexual activity and sexID stratification

## Estimate

```{r}
library(reshape2)
gsisa_rep <- all_sex %>%
    group_by(r_sex, r_sexid, r_sexact) %>%
    summarize(het = weighted.mean(hetnonew, w = total_wt),
              sam = weighted.mean(samnonew, w = total_wt)) %>%
    mutate(m = ifelse(r_sex == "m", sam, het),
           w = ifelse(r_sex == "w", sam, het)) %>%
    select(r_sex, r_sexid, r_sexact, m, w) %>%
    melt(id.vars = c('r_sex', 'r_sexid', 'r_sexact'),
         variable.name = "rp_sex",
         value.name = "partners") %>%
    # add proportions
    rowwise() %>%
    mutate(prop = get_proportion(all_sex, q_sex = r_sex,
                                q_sexid = r_sexid, q_sexact = r_sexact),
           n_partners = prop * partners) %>%
    ungroup() %>%
  arrange(r_sex, r_sexid, r_sexact)
format_table(gsisa_rep)
```


```{r}
gsisa_offered <- gsisa_rep %>%
  group_by(rp_sex, r_sex) %>%
  # define prop_to_r as the the proportion of all partners offered to r that come from rp
  mutate(prop_to_r = n_partners / sum(n_partners)) %>%
  select(r_sex,
         rp_sex,
         rp_sexid = r_sexid,
         rp_sexact = r_sexact,
         prop_to_r) %>%
  arrange(r_sex, rp_sex, rp_sexid, rp_sexact)
format_table(gsisa_offered)
```

Note that most of the partners that are 'offered' from men to men from men come from high sexual-activity gay men, and the proportion of all partnerships offered from men to men sums to 1.

```{r}
fmtm <- gsisa_offered %>%
  filter(r_sex == "m" & rp_sex == "m") %>%
  mutate(rp_demo = paste(rp_sex, rp_sexid, rp_sexact, sep="_"))
ggplot(fmtm) +
  geom_bar(aes(x = rp_demo, y = prop_to_r), stat = "identity") +
  ylab("Proportion of partnerships offered to men from men") +
  theme_bw()
```

Make partner distribution

## Distribute

```{r}
make_partner_dist_gsisa <- function(survey_df, dist_df){
    joined <- 
    return(joined)
}

gsisa_partner_dist <- left_join(gsisa_rep, gsisa_offered, by = c("r_sex", "rp_sex")) %>%
        mutate(d_partners = partners * prop_to_r,
               n_d_partners = d_partners * prop) %>%
        mutate(r_demo = paste(r_sex, r_sexid, r_sexact, sep="_"),
               rp_demo = paste(rp_sex, rp_sexid, rp_sexact, sep="_")) %>%
  select(r_demo, rp_demo, prop, d_partners)

format_table(head(gsisa_partner_dist, n = 10))
```

## Balance

```{r}
balance_gsisa <- function(partner_dist_gsisa, theta){
    bidi <- partner_dist_gsisa %>%
        inner_join(partner_dist_gsisa,
                   by = c("r_demo" = "rp_demo",
                          "rp_demo" = "r_demo"),
                   suffix = c('.r', '.rp'))
    
    b_bidi <- bidi %>%
        mutate(np_r = d_partners.r * prop.r,
               np_rp = d_partners.rp * prop.rp) %>%
        mutate(imbalance = np_r / np_rp,
               corrected_r = d_partners.r / imbalance^(1 - theta),
               corrected_rp = d_partners.rp * imbalance^theta,
               cnr = corrected_r * prop.r,
               cnrp = corrected_rp * prop.rp) %>%
        select(r_demo,
               prop.r,
               rp_demo,
               prop.rp,
               d_partners.r,
               corrected_r,
               d_partners.rp,
               corrected_rp,
               cnr,
               cnrp)
    return(b_bidi)
}

b_gsisa_partner_dist <- balance_gsisa(gsisa_partner_dist, 0.5)

format_table(head(b_gsisa_partner_dist, n = 10))

library(stringr)
ggplot(rbind(b_gsisa_partner_dist)) +
    geom_tile(aes(x = r_demo, y = rp_demo, fill=corrected_r), color="black", size = 0.2) +
    scale_fill_gradient2(name="Annual Number\nof New Partnerships",
                         low = "white",
                         mid = "blue2",
                         high = "black",
                         midpoint = 3,
                         na.value = "white") +
    labs(x = "Group of Partner 1",
         y = "Group of Partner 2") +
    theme_minimal(base_size = 14) + coord_fixed() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
# ggsave("plots/gsisa_partner_dist.png", width = 7, height=6, units = 'in', dpi =500)
```

Check that this worked:

```{r}
all(with(b_gsisa_partner_dist, abs(cnr - cnrp) < .Machine$double.eps), na.rm = TRUE)
```

### Compare
```{r}
sexid_nos_by_grp <- b_gsisa_partner_dist %>%
  mutate(rp_sex = str_split(rp_demo, "_", simplify = TRUE)[, 1]) %>%
  group_by(r_demo, rp_sex) %>%
  summarize(bal_p_pc = sum(corrected_r, na.rm = TRUE))

sexid_nos_compare <- gsisa_rep %>%
  mutate(r_demo = paste(r_sex, r_sexid, r_sexact, sep = "_")) %>%
  select(r_demo, rp_sex, orig_p_pc = partners) %>%
  left_join(sexid_nos_by_grp, by = c("r_demo", "rp_sex"))
format_table(sexid_nos_compare)
```

# Appendix

## Code Variable Names
Throughout this document, we use the following nomenclature:

| Name         | Definition |
|--------------|------------------------------------------------------------------|
|`r_sex` | the sex of the survey respondent |
|`r_sexid`    | the self-reported sexual identity of the survey respondent |
|`r_sexact`     | the sexual activity group of the survey respondent, based on their total number of partners |
| `prop` | the proportion of the total population with the designated sex, sexid, and/or sexact. This may be suffixed by `r_rp` or `rp_r` |
|`rp_sex`       | the sex of the sex partners (respondents' partners). This is reported by the survey respondents. |
|`rp_sexid`         | the sexual identity of the sex partners of the respondents. This will be estimated using the proportionality assumption |
|`rp_sexact` | the sexual activity group of the sex partners of the respondents. This will be estimated using the proportionality assumption. |
|`partners` | the per-person number of new sex partners that the `r` group reported with people of sex `rp_sex` |
|`n_partners` | the total number of new sex partners that the `r` group reported with people of sex `rp_sex` |
|`partners.r_rp`| After combining respondents and hypothetical respondents' partners, the per-person number of partners that `r` has with `rp` |
|`partners.rp_r`| The hypothetical per-person number of partnerships that `rp` has with `r` |
|`prop_to_r` | The proportion of all partnerships offered to `r` that come from `rp` |
|`d_partners`, `d_partners.r_rp`, `d_partners.rp_r`| The per-person number of partnerships, distributed over sexids and/or sex activity groups. Either as reported, calculated from `r` to `rp`, or calculated from `rp` to `r`, respectively. |
|`n_partners*`, `n_d_partners*` | The total number of partnerships, i.e., the product of the partnership measure and the proportion in that group |
|`corrected_r`, `corrected_rp` | Per-person number of partnerships corrected for balancing, from the perspective of `r` and `rp` groups, respectively.
|`cnr`, `cnrp` | The total number of partnerships from `r` to `rp` and vice versa. Used to check that the balancing worked |

